\textbf{Microbiota predict \emph{Clostridium difficile} severity in germ-free mice colonized with human feces}

### 2016-12-05 submission/Manuscript.Rmd
got citations working, not done yet

---
title: "**Microbiota predict _Clostridium difficile_ severity in germ-free mice colonized with human feces**"
output:
  pdf_document:
    includes:
      in_header: header.tex
    keep_tex: yes
csl: mbio.csl #Get themes at https://github.com/citation-style-language/styles
fontsize: 11pt
geometry: margin=1.0in
---

@@

### Abstract
_Clostridium difficile_ causes diarrheal disease when it successfully colonizes a disrupted gut microbial community. Current mouse models to study _C. difficile_ infection (CDI) rely on pre-treatment with antibiotics to alter the mouse microbiome prior to inoculation. This model does not allow for analysis of human-associated microbial community members that modulate _C. difficile_ colonization and expansion. To study human-associated microbes in the context of CDI, we inoculated germ-free C57BL/6 mice with one of 16 human fecal samples from diarrheal or healthy donors and challenged with C. difficile 14 days later. Five unique donor-mice combinations resulted in severe CDI while the remaining 11 only experienced mild disease. Both healthy and diarrheal donors were susceptible to colonization and severe symptoms of CDI. To determine if specific microbes were associated with disease severity outcomes, we built a classification Random Forest machine learning model based on relative abundance data of the communities prior to infection. The model identified a number of bacterial populations associated with the development of severe CDI, including _Bacilliales, Ruminococcaceae, Ruminococcus, Staphylococcus, Streptococcus and Bacteriodetes_. Additionally, a regression model accurately predicted colonization levels of _C. difficile_ at one to ten days post-infection. This model explained 99% of the variance in the number of CFU isolated from mouse stool. Members of _Lachnospiraceae, Parabacteroides, Bacteroidales, Bacteroidetes, Porphyromonadaceae_ and unclassified _Bacteria_ families were predictive of future _C. difficile_ colonization levels. Finally, challenging these mice with different strains of _C. difficile_ revealed that susceptible human-associated microbial communities were prone to severe disease independent of strain type. Taken together these results suggest that human-associated microbial communities can be recapitulated in germ-free mice and used to characterize dynamics of CDI. Because both healthy and diarrheal patients were susceptible to severe CDI, machine-learning models are useful to identify bacterial populations that allow colonization and contribute to the development of _C. difficile_ associated disease in humans.

@@

### Introduction
_Clostridium difficile_ is an opportunistic pathogen of the human lower gastrointestinal tract. _C. difficile_ forms spores that can persist on abiotic surfaces and are not readily killed by ethanol-based hand-sanitizers, putting hospital patients particularly at risk. Indeed, ~12% of hospital acquired infections in the United States are due to _C. difficile_ and result in up to 15,000 deaths annually (2). Disruption of the native microbial community is the most common risk factor for development of _C. difficile_ infection (CDI) (1). Antibiotic use and inflammatory bowel diseases are associated with loss of colonization resistance to _C. difficile_ through the loss of potentially protective bacterial families such as _Barnesiella_ and _Lachnospiraceae_ in both mouse models and human association experiments (cite alyx 2014 and vincent 2013). The composition of the community is clearly important for the acquisition, resistance and treatment of _C. difficile_ infection, as giving patients a healthy fecal microbiome transplant is the most effective treatment for this disease (cite Anna 2014). The precise mechanisms of colonization resistance and restoration of a healthy community are yet to be discovered.

>>>>

---
title: "**Microbiota predict _Clostridium difficile_ severity in germ-free mice colonized with human feces**"
bibliography: references.bib
output:
  pdf_document:
    includes:
      in_header: header.tex
    keep_tex: yes
csl: plos.csl #Get themes at https://github.com/citation-style-language/styles
fontsize: 11pt
geometry: margin=1.0in
---

@@

### Abstract
_Clostridium difficile_ causes diarrheal disease when it successfully colonizes a disrupted gut microbial community. Current mouse models to study _C. difficile_ infection (CDI) rely on pre-treatment with antibiotics to alter the mouse microbiome prior to inoculation. This model does not allow for analysis of human-associated microbial community members that modulate _C. difficile_ colonization and expansion. To study human-associated microbes in the context of CDI, we inoculated germ-free C57BL/6 mice with one of 16 human fecal samples from diarrheal or healthy donors and challenged with _C. difficile_ 14 days later. Five unique donor-mice combinations resulted in severe CDI while the remaining 11 only experienced mild disease. Both healthy and diarrheal donors were susceptible to colonization and severe symptoms of CDI. To determine if specific microbes were associated with disease severity outcomes, we built a classification Random Forest machine learning model based on relative abundance data of the communities prior to infection. The model identified a number of bacterial populations associated with the development of severe CDI, including _Bacilliales, Ruminococcaceae, Ruminococcus, Staphylococcus, Streptococcus and Bacteriodetes_. Additionally, a regression model accurately predicted colonization levels of _C. difficile_ at one to ten days post-infection. This model explained 99% of the variance in the number of CFU isolated from mouse stool. Members of _Lachnospiraceae, Parabacteroides, Bacteroidales, Bacteroidetes, Porphyromonadaceae_ and unclassified _Bacteria_ families were predictive of future _C. difficile_ colonization levels. Finally, challenging these mice with different strains of _C. difficile_ revealed that susceptible human-associated microbial communities were prone to severe disease independent of strain type. Taken together these results suggest that human-associated microbial communities can be recapitulated in germ-free mice and used to characterize dynamics of CDI. Because both healthy and diarrheal patients were susceptible to severe CDI, machine-learning models are useful to identify bacterial populations that allow colonization and contribute to the development of _C. difficile_ associated disease in humans.

@@

### Introduction
_Clostridium difficile_ is an opportunistic pathogen of the human lower gastrointestinal tract. _C. difficile_ forms spores that can persist on abiotic surfaces and are not readily killed by ethanol-based hand-sanitizers, putting hospital patients particularly at risk. Indeed, ~12% of hospital acquired infections in the United States are due to _C. difficile_ and result in up to 15,000 deaths annually [@Lessa2015]. Disruption of the native microbial community is the most common risk factor for development of _C. difficile_ infection (CDI) [@Britton2012]. Antibiotic use and inflammatory bowel diseases are associated with loss of colonization resistance to _C. difficile_ through the loss of potentially protective bacterial families such as _Barnesiella_ and _Lachnospiraceae_ in both mouse models and human association experiments [@Schubert2014;@Vincent2013]. The composition of the community is clearly important for the acquisition, resistance and treatment of _C. difficile_ infection, as giving patients a healthy fecal microbiome transplant is the most effective treatment for this disease [@Seekatz2014]. The precise mechanisms of colonization resistance and restoration of a healthy community are yet to be discovered.

### 2016-12-05 submission/Manuscript.tex
got citations working, not done yet

\emph{Clostridium difficile} causes diarrheal disease when it
successfully colonizes a disrupted gut microbial community. Current
mouse models to study \emph{C. difficile} infection (CDI) rely on
pre-treatment with antibiotics to alter the mouse microbiome prior to
inoculation. This model does not allow for analysis of human-associated
microbial community members that modulate \emph{C. difficile}
colonization and expansion. To study human-associated microbes in the
context of CDI, we inoculated germ-free C57BL/6 mice with one of 16
human fecal samples from diarrheal or healthy donors and challenged with
C. difficile 14 days later. Five unique donor-mice combinations resulted
in severe CDI while the remaining 11 only experienced mild disease. Both
healthy and diarrheal donors were susceptible to colonization and severe
symptoms of CDI. To determine if specific microbes were associated with
disease severity outcomes, we built a classification Random Forest
machine learning model based on relative abundance data of the
communities prior to infection. The model identified a number of
bacterial populations associated with the development of severe CDI,
including \emph{Bacilliales, Ruminococcaceae, Ruminococcus,
Staphylococcus, Streptococcus and Bacteriodetes}. Additionally, a
regression model accurately predicted colonization levels of \emph{C.
difficile} at one to ten days post-infection. This model explained 99\%
of the variance in the number of CFU isolated from mouse stool. Members
of \emph{Lachnospiraceae, Parabacteroides, Bacteroidales, Bacteroidetes,
Porphyromonadaceae} and unclassified \emph{Bacteria} families were
predictive of future \emph{C. difficile} colonization levels. Finally,
challenging these mice with different strains of \emph{C. difficile}
revealed that susceptible human-associated microbial communities were
prone to severe disease independent of strain type. Taken together these
results suggest that human-associated microbial communities can be
recapitulated in germ-free mice and used to characterize dynamics of
CDI. Because both healthy and diarrheal patients were susceptible to
severe CDI, machine-learning models are useful to identify bacterial
populations that allow colonization and contribute to the development of
\emph{C. difficile} associated disease in humans.

@@

\emph{Clostridium difficile} is an opportunistic pathogen of the human
lower gastrointestinal tract. \emph{C. difficile} forms spores that can
persist on abiotic surfaces and are not readily killed by ethanol-based
hand-sanitizers, putting hospital patients particularly at risk. Indeed,
\textasciitilde{}12\% of hospital acquired infections in the United
States are due to \emph{C. difficile} and result in up to 15,000 deaths
annually (2). Disruption of the native microbial community is the most
common risk factor for development of \emph{C. difficile} infection
(CDI) (1). Antibiotic use and inflammatory bowel diseases are associated
with loss of colonization resistance to \emph{C. difficile} through the
loss of potentially protective bacterial families such as
\emph{Barnesiella} and \emph{Lachnospiraceae} in both mouse models and
human association experiments (cite alyx 2014 and vincent 2013). The
composition of the community is clearly important for the acquisition,
resistance and treatment of \emph{C. difficile} infection, as giving
patients a healthy fecal microbiome transplant is the most effective
treatment for this disease (cite Anna 2014). The precise mechanisms of
colonization resistance and restoration of a healthy community are yet
to be discovered.

@@

\textbf{Nick will be updating the next two sections} \textbf{Microbes
present in the gut prior to infection are predictive of \emph{C.
difficile} colonization levels} Previous work in our group has
demonstrated that the microbes present in the mouse microbiome prior to
\emph{C. difficile} challenge can predict future colonization levels
(cite Alyx). To determine if the human-associated microbiome in the
mouse could similarly predict \emph{C. difficile} colonization we
employed a Random Forest machine learning algorithm similar to the one
used in previous studies. Relative abundance of OTUs with greater than
1\% abundance at day 0 were used as input to the model. We found that
the model explained X\% of the variance in colonization levels of
\emph{C. difficile} (Fig 3A). By refining the model to include only the
top X predictive OTUs, the model explained X\% of the variance. Partial
dependency plots of the top 6 OTUs predictive of \emph{C. difficile}
colonization reveal that colonization is positively associated with
\emph{Bacteriodetes}, \emph{Parabacteriodes} and
\emph{Porphyromonadaceae} (really?) (Fig 3B). Colonization was
negatively associated with the presence of \emph{Lachnospiraceae} (Fig
3B) though none of the human-associated mouse communities were
completely resistant to colonization by \emph{C. difficile} (Fig 2A,
3B). These results confirm that as in mouse microbiome studies, the
human-associated gut community prior to challenge is predictive of
\emph{C. difficile} colonization levels.

@@

\textbf{Susceptibility to infection is community-dependent.} \emph{C.
difficile} strains are known to vary in disease phenotypes based both on
pathogenicity factors of the strain and host microbiome composition
(cite someone). To test the susceptibility of human-derived microbiomes
to \emph{C. difficile} infection, two donor-mouse microbiomes that were
previously shown to be resistant to \emph{C. difficile} strain 431
induced weight loss (Fig 2B) (DA00396, DA00430) and one that was
suscpetible to severe disease (DA00578) were independently challenged
with one of three unique \emph{C. difficile} strains (table or reference
for toxin prod, etc). As done previously, \emph{C. difficile} stool CFU
was enumerated daily and mouse weight-loss tracked. Infection of mice
with resistant communities with virulent strain 458 or the weaker strain
299 resulted in colonization to levels near that of 431 (Fig 5A) and
these mice did not experience severe weight loss or show clinical signs
of disease (Fig 5B). In contrast, infection of mice harboring
microbiomes susceptible to severe disease (DA00578, Fig 1B) with strain
395 resulted in colonization and severe weight loss by two days
post-infection (Fig 5A, 5B). Strain 395 is similarly virulent as strain
431, and shows a similar infection phenotype in these mice. These
preliminary results suggest that human-associated microbiomes may be
more predictive of disease severity than the \emph{C. difficile} strain.

@@

Study of human-associated microbiomes and their interactions with
\emph{C. difficile} has been limited by the lack of an appropriate model
and available samples. Here we colonized 16 groups of mice with human
donor stool from patients with different clinical profiles or healthy
volunteers. This created a diverse set of human-associated microbiomes
in the mice that allowed for testing \emph{C. difficile} infection
dynamics and community interactions. Our results showed that both
healthy and diarrheal-associated human communities were susceptible to
severe disease. Additionally, our results confirmed our and others
previous studies that proved that resistance to \emph{C. difficile}
disease is dependent on the entire microbial community and not just one
or two species (cite alyx). Further, we were able to learn
classification models that identified specific OTUs within susceptible
communities that predispose mice to infection. By subsequently infecting
susceptible or resistant communities with varying strains of \emph{C.
difficile} we confirmed that the initial state of the human-associated
microbiome is more predictive of disease severity than the strain used.

Here we were able to colonize germ-free mice with a variety of human
stool communities from diarrheal or healthy people. After allowing these
communities to colonize the mice for 2 weeks, they were challenged with
\emph{C. difficile}. It is important to note that the mouse gut
environment is different from that of the human, and as such exerts a
selective pressure on the human stool microbial community that
stabilizes over the 2 weeks. Thus these mice cannot be considered
``humanized'' but rather containing human-associated microbes.
Interestingly, clinical status was not associated with proclivity to
severe \emph{C. difficile} infection as some healthy donor mice were
susceptible. A surprising finding was that none of the human-associated
mouse communities were completely resistant to colonization by \emph{C.
difficile}. This is in contrast to numerous reports of mouse microbiome
studies (cite) and the knowledge that in general healthy humans without
gut perturbations are resistant to \emph{C. difficile} infection (cite).
On the other hand, it is known that humans, especially health care
employees, can be asymptomatic carriers of \emph{C. difficile}
(Furyura-Kanamari). This may explain the fact that the majority of our
groups of mice were colonized by \emph{C. difficile} but did not display
outward signs of illness or infection-associated weight loss (Fig 2).

@@

Finally, our pilot experiments testing the efficiency of infection with
different \emph{C. difficile} strains have allowed for testing community
versus pathogen factors of infection. Our results are consistent with
previous reports that the composition of the gut microbiome determines
if an individual is successfully infected or not (cite Alyx and others).
\emph{C. difficile} strains are known to vary in pathogenecity,
primarily via the amount of toxin that is produced during infection
(cite). In these pilot experiments, we challenged a susceptible
community with another pathogenic strain of \emph{C. difficile} to
confirm that the susceptibility lied in the community and not in the
specific challenge strain. In contrast, we exposed a
resistant/asymptomatic community to a relatively weak strain of \emph{C.
difficile} and found that the community could be colonized by this one
as well-- but not cause disease. This suggests that this colony is prone
to colonization by any \emph{C. difficile} strain. Further experiments
will test if infecting susceptible communities with weak strains allows
the mice to survive. It has been recently reported that carriage of a
non-toxigenic strain of \emph{C. difficile} can prevent additional
colonization by a pathogenic strain, suggesting some role of
niche-filling or immunity this provides (cite cite). It would be
interesting to see if we could make a susceptible community resistant to
infection and disease by a pathogenic \emph{C. difficle} strain in our
human-associated microbiome mice. Further work with human-associated
communities in mice and mathematical modeling will allow for uncovering
of complex human-associated disease processes and dynamics as they
relate to human disease.

@@

\textbf{\emph{Clostridium difficile} and sample processing} For all
experiments, 100 spores of \emph{Clostridium difficile} were used to
orally infect mice as described previously (cite alyx). The \emph{C.
difficile} strains used were acquired from spore stocks of (WHERE) and
comprised of strains 431, 395, 299 and 468 (cite?). After gavage, the
remaining inoculum was diluted and plated to confirm infection dose.
Bacteria were cultured anaerobically on TCCFA plates and incubated at
37C. During infection, two fresh stool pellets were collected from each
mouse daily. One pellet was resuspended in anaerobic phosphate-buffered
saline and diluted and plated on TCCFA plates to enumerate \emph{C.
difficile} CFU. The other pellet was frozen at -80C until the end of the
experiment.

@@

We would like to thank members of the Schloss lab and Jhansi Leslie for
analysis advice and the Microbial Systems Laboratory at the University
of Michigan for sequencing assistance.

@@

\subsection{References}\label{references}

>>>>

\emph{Clostridium difficile} causes diarrheal disease when it
successfully colonizes a disrupted gut microbial community. Current
mouse models to study \emph{C. difficile} infection (CDI) rely on
pre-treatment with antibiotics to alter the mouse microbiome prior to
inoculation. This model does not allow for analysis of human-associated
microbial community members that modulate \emph{C. difficile}
colonization and expansion. To study human-associated microbes in the
context of CDI, we inoculated germ-free C57BL/6 mice with one of 16
human fecal samples from diarrheal or healthy donors and challenged with
\emph{C. difficile} 14 days later. Five unique donor-mice combinations
resulted in severe CDI while the remaining 11 only experienced mild
disease. Both healthy and diarrheal donors were susceptible to
colonization and severe symptoms of CDI. To determine if specific
microbes were associated with disease severity outcomes, we built a
classification Random Forest machine learning model based on relative
abundance data of the communities prior to infection. The model
identified a number of bacterial populations associated with the
development of severe CDI, including \emph{Bacilliales, Ruminococcaceae,
Ruminococcus, Staphylococcus, Streptococcus and Bacteriodetes}.
Additionally, a regression model accurately predicted colonization
levels of \emph{C. difficile} at one to ten days post-infection. This
model explained 99\% of the variance in the number of CFU isolated from
mouse stool. Members of \emph{Lachnospiraceae, Parabacteroides,
Bacteroidales, Bacteroidetes, Porphyromonadaceae} and unclassified
\emph{Bacteria} families were predictive of future \emph{C. difficile}
colonization levels. Finally, challenging these mice with different
strains of \emph{C. difficile} revealed that susceptible
human-associated microbial communities were prone to severe disease
independent of strain type. Taken together these results suggest that
human-associated microbial communities can be recapitulated in germ-free
mice and used to characterize dynamics of CDI. Because both healthy and
diarrheal patients were susceptible to severe CDI, machine-learning
models are useful to identify bacterial populations that allow
colonization and contribute to the development of \emph{C. difficile}
associated disease in humans.

@@

\emph{Clostridium difficile} is an opportunistic pathogen of the human
lower gastrointestinal tract. \emph{C. difficile} forms spores that can
persist on abiotic surfaces and are not readily killed by ethanol-based
hand-sanitizers, putting hospital patients particularly at risk. Indeed,
\textasciitilde{}12\% of hospital acquired infections in the United
States are due to \emph{C. difficile} and result in up to 15,000 deaths
annually {[}1{]}. Disruption of the native microbial community is the
most common risk factor for development of \emph{C. difficile} infection
(CDI) {[}2{]}. Antibiotic use and inflammatory bowel diseases are
associated with loss of colonization resistance to \emph{C. difficile}
through the loss of potentially protective bacterial families such as
\emph{Barnesiella} and \emph{Lachnospiraceae} in both mouse models and
human association experiments {[}3,4{]}. The composition of the
community is clearly important for the acquisition, resistance and
treatment of \emph{C. difficile} infection, as giving patients a healthy
fecal microbiome transplant is the most effective treatment for this
disease (cite Anna 2014). The precise mechanisms of colonization
resistance and restoration of a healthy community are yet to be
discovered.

@@

\textbf{Nick will be updating the next two sections} \textbf{Microbes
present in the gut prior to infection are predictive of \emph{C.
difficile} colonization levels} Previous work in our group has
demonstrated that the microbes present in the mouse microbiome prior to
\emph{C. difficile} challenge can predict future colonization levels
(cite Alyx). To determine if the human-associated microbiome in the
mouse could similarly predict \emph{C. difficile} colonization we
employed a Random Forest machine learning algorithm similar to the one
used in previous studies. Relative abundance of OTUs with greater than
1\% abundance at day 0 were used as input to the model. We found that
the model explained X\% of the variance in colonization levels of
\emph{C. difficile} (Fig 3A). By refining the model to include only the
top X predictive OTUs, the model explained X\% of the variance. Partial
dependency plots of the top X OTUs predictive of \emph{C. difficile}
colonization reveal that colonization is positively associated with
\emph{Bacteriodetes}, \emph{Parabacteriodes} and
\emph{Porphyromonadaceae} (really?) (Fig 3B). Colonization was
negatively associated with the presence of \emph{Lachnospiraceae} (Fig
3B) though none of the human-associated mouse communities were
completely resistant to colonization by \emph{C. difficile} (Fig 2A,
3B). These results confirm that as in mouse microbiome studies, the
human-associated gut community prior to challenge is predictive of
\emph{C. difficile} colonization levels.

@@

\textbf{Susceptibility to infection is community-dependent.} \emph{C.
difficile} strains are known to vary in disease phenotypes based both on
pathogenicity factors of the strain and host microbiome composition
(cite someone). To test the susceptibility of human-derived microbiomes
to \emph{C. difficile} infection, two donor-mouse microbiomes that were
previously shown to be resistant to \emph{C. difficile} strain 430
induced weight loss (Fig 2B) (DA00396, DA00430-wait isnt this a control)
and one that was susceptible to severe disease (DA00578) were
independently challenged with one of three unique \emph{C. difficile}
strains (\emph{table or reference for toxin prod}, etc). As done
previously, \emph{C. difficile} stool CFU was enumerated daily and mouse
weight-loss tracked. Infection of resistant mice with virulent strain
458 or the weaker strain 299 resulted in colonization to levels near
that of 431 (Fig 5A). Accordingly, these mice did not experience severe
weight loss or show clinical signs of disease (Fig 5B). In contrast,
infection of susceptible mice (DA00578, Fig 1B) with pathogenic strain
395 resulted in colonization and severe weight loss by two days
post-infection (Fig 5A, 5B). Strain 395 is similarly virulent as strain
431, and shows a similar infection phenotype in these mice. These
preliminary results suggest that human-associated microbiomes may be
more predictive of disease severity than the \emph{C. difficile} strain.

@@

Study of human-associated microbiomes and their interactions with
\emph{C. difficile} has been limited by the lack of an appropriate model
and available samples. Here we colonized 16 groups of mice with human
donor stool from patients with different clinical profiles or healthy
volunteers. This created a diverse set of human-associated mouse
microbiomes that allowed for testing \emph{C. difficile} infection
dynamics and community interactions. Our results showed that both
healthy and diarrheal-associated human communities were susceptible to
severe disease. Additionally, our results confirmed previous findings
that resistance to \emph{C. difficile} disease is dependent on the
entire microbial community and not just one or two taxa (cite alyx).
Further, we were able to learn classification models that identified
specific OTUs within susceptible communities that predispose mice to
infection. By subsequently infecting susceptible or resistant
communities with varying strains of \emph{C. difficile} we confirmed
that the initial state of the human-associated microbiome is more
predictive of disease severity than the strain used.

Here we colonized germ-free mice with a variety of human stool
communities and allowed these communities to colonize the mice for 2
weeks, prior to \emph{C. difficile} challenge. It is important to note
that the mouse gut environment is different from that of the human, and
as such exerts a selective pressure on the human stool microbial
community that stabilizes over the 2 weeks. Thus these mice cannot be
considered ``humanized'' but rather colonized by human-associated
microbes. Interestingly, clinical status was not associated with
proclivity to severe \emph{C. difficile} infection as some healthy donor
mice were susceptible. A surprising finding was that none of the
human-associated mouse communities were completely resistant to
colonization by \emph{C. difficile}. This is in contrast to numerous
reports of mouse microbiome studies (cite) and the knowledge that in
general healthy humans without gut perturbations are resistant to
\emph{C. difficile} infection (cite). On the other hand, it is known
that humans, especially health care employees, can be asymptomatic
carriers of \emph{C. difficile} (Furyura-Kanamari). This may explain the
fact that the majority of our groups of mice were colonized by \emph{C.
difficile} but did not display outward signs of illness or
infection-associated weight loss (Fig 2).

@@

Finally, our pilot experiments testing the efficiency of infection with
different \emph{C. difficile} strains have allowed for testing community
versus pathogen factors of infection. Our results are consistent with
previous reports that the composition of the gut microbiome determines
if an individual is successfully infected or not (cite Alyx and others).
\emph{C. difficile} strains are known to vary in pathogenicity,
primarily via the amount of toxin that is produced during infection
(cite). To confirm that disease risk is dependent on the microbiome, we
challenged susceptibile mice with an additional pathogenic strain of
\emph{C. difficile} and found that the mice indeed fell ill. In
contrast, we exposed an asymptomatic community to a relatively weak
strain of \emph{C. difficile} and found that the community could be
colonized by this one as well-- but not cause disease. This suggests
that this particular colony is prone to colonization by any \emph{C.
difficile} strain. Further experiments will test if infecting
susceptible communities with weak strains allows the mice to survive. It
has been recently reported that carriage of a non-toxigenic strain of
\emph{C. difficile} can prevent additional colonization by a pathogenic
strain, suggesting a role of niche-filling or immunity this
pre-colonization provides (cite cite). One interesting experiment would
be to recapitulate this finding in human-associated mouse microbiomes by
making a susceptible community resistant to infection. Further work with
human-associated communities in mice and mathematical modeling will
allow for uncovering of complex human-associated disease processes and
dynamics as they relate to human \emph{C. difficile} infection.

@@

\textbf{\emph{Clostridium difficile} and sample processing} For all
experiments, 100 spores of \emph{Clostridium difficile} were used to
orally infect mice as described previously (cite alyx). The \emph{C.
difficile} strains used were acquired from spore stocks of (WHERE) and
comprised of strains 431, 395, 299 and 468 (cite?). After gavage, the
remaining inoculum was diluted and plated to confirm infection dose.
Bacteria were cultured anaerobically on TCCFA plates and incubated at
37C (cite plate recipe). During infection, two fresh stool pellets were
collected from each mouse daily. One pellet was resuspended in anaerobic
phosphate-buffered saline and diluted and plated on TCCFA plates to
enumerate \emph{C. difficile} CFU. The other pellet was frozen at -80C
until the end of the experiment.

@@

We would like to thank members of the Schloss lab and Jhansi Leslie for
analysis advice. We would also like to thank the Microbial Systems
Laboratory at the University of Michigan for sequencing assistance.

@@

\subsection*{References}\label{references}
\addcontentsline{toc}{subsection}{References}

1. Lessa FC, Mu Y, Bamberg WM, Beldavs ZG, Dumyati GK, Dunn JR, et al.
Burden ofClostridium difficileInfection in the united states. New
England Journal of Medicine. New England Journal of Medicine (NEJM/MMS);
2015;372: 825--834.
doi:\href{http://dx.doi.org/10.1056/nejmoa1408913}{10.1056/nejmoa1408913}

2. Britton RA, Young VB. Interaction between the intestinal microbiota
and host in clostridium difficile colonization resistance. Trends in
Microbiology. Elsevier BV; 2012;20: 313--319.
doi:\href{http://dx.doi.org/10.1016/j.tim.2012.04.001}{10.1016/j.tim.2012.04.001}

3. Schubert AM, Rogers MAM, Ring C, Mogle J, Petrosino JP, Young VB, et
al. Microbiome data distinguish patients with clostridium difficile
infection and non-c. difficile-associated diarrhea from healthy
controls. mBio. American Society for Microbiology; 2014;5:
e01021--14--e01021--14.
doi:\href{http://dx.doi.org/10.1128/mbio.01021-14}{10.1128/mbio.01021-14}

4. Vincent C, Stephens DA, Loo VG, Edens TJ, Behr MA, Dewar K, et al.
Reductions in intestinal clostridiales precede the development of
nosocomial clostridium difficile infection. Microbiome. Springer Nature;
2013;1: 18.
doi:\href{http://dx.doi.org/10.1186/2049-2618-1-18}{10.1186/2049-2618-1-18}
